### 2019-12-04 cards_persistent_analysis.Rmd
Some more analysis of what's going on, and starting to derive what selection criteria might be for model eval.

---
title: "Cards persistent analysis"
author: "Andrew Lampinen"
output: html_document
---

```{r}
library(tidyverse)
```

# data loading

```{r}
parent_dir = "."
subdirs = c("results_for_humans_with_library_2", "results_for_humans_with_library_3",
            #"results_for_humans_with_library_persistent",  "results_for_humans_with_library_persistent_2", "results_for_humans_with_library_persistent_3",  "results_for_humans_with_library_persistent_4", 
            "results_for_humans_with_library_persistent_5",  "results_for_humans_with_library_persistent_6",
            "results_for_humans_with_library_persistent_7", 
            "results_for_humans_with_library_persistent_8", 
            "results_for_humans_with_library_persistent_9"
            )
num_runs = 3
```

```{r}
read_config = function(config_file) { 
  config = read_delim(config_file, delim="\n") %>%
    separate(`key, value`, c("key", "value"), sep=",", extra="merge") %>%
    spread(key, value) %>%
    mutate_at(c("base_train_tasks", "base_eval_tasks", "meta_class_train_tasks", "meta_class_eval_tasks", "meta_map_train_tasks", "meta_map_eval_tasks"), function(x) {
      x = gsub("\\\"|[][]| |\'", "", x)
      return(str_split(x, ","))
    } )
}
```

```{r}
load_d = function(results_dir, result_subdirs, num_runs, file_type) {
  d = data.frame()
  for (run_i in 0:(num_runs-1)) {
    for (result_subdir in result_subdirs) {
      filename = sprintf("%s/%s/run%i_%s.csv", results_dir, result_subdir, run_i, file_type)
      print(filename)
      if (!file.exists(filename)) {
        print(paste("skipping ", filename, sep=""))
        next
      }
      if (grepl("config", file_type)) {
        this_d = read_config(filename)
      } else {
        this_d = read.csv(filename, check.names=F, header=T) 
        names(this_d) <- make.unique(names(this_d))

      }
      this_d = this_d %>%
        mutate(run = run_i,
               run_type = result_subdir)
      d = d %>%
        bind_rows(this_d)
    }

  }
  return(d)
}
```

```{r}
config_d = load_d(parent_dir, subdirs, num_runs, "run_config")
loss_d = load_d(parent_dir, subdirs, num_runs, "losses")
meta_true_d = load_d(parent_dir, subdirs, num_runs, "meta_true_losses")
```

# some manipulation

```{r}
loss_d = loss_d %>%
  filter(epoch %% 100 == 0) %>%
  gather(task, loss, -epoch, -run, -run_type) %>%
  mutate(meta = grepl("is|toggle", task)) %>%
  filter(!meta, grepl("rewards", task)) %>%
  mutate(task = sub("_rewards", "", task),
         train_or_eval = ifelse(grepl("straight_flush_l_1", task), "eval", "train"))

```

```{r}
meta_true_d = meta_true_d %>%
  filter(epoch %% 100 == 0) %>%
  gather(task, loss, -epoch, -run, -run_type) %>%
  separate(task, c("meta_task", "mapping_toe", "base_task_toe", "source", "target"), ":|->")

```

# basic plots
```{r}
theme_set(theme_bw())
```

```{r}
ggplot(loss_d,
       aes(x=epoch, y=loss, color=train_or_eval)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run_type ~ meta)
```

```{r}
ggplot(meta_true_d %>% 
         filter(meta_task == "toggle_losers"),
       aes(x=epoch, y=loss, color=base_task_toe)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run ~ run_type)
```

```{r}
ggplot(meta_true_d %>% 
         filter(meta_task == "toggle_losers",
                source == "game_straight_flush_l_0_bv_0_sr_0"),
       aes(x=epoch, y=loss, color=base_task_toe)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run ~ run_type)
```

```{r}
ggplot(meta_true_d %>%
         filter(meta_task=="toggle_losers") %>%
         group_by(run_type, run, epoch, base_task_toe) %>%
         summarize(mean_loss=mean(loss)) %>%
         ungroup() %>%
         spread(base_task_toe, mean_loss),
       aes(x=example_is_train, y=example_is_eval, alpha=epoch)) +
  geom_point() +
  facet_grid(run ~ run_type)
```

```{r}
selected_meta_true_d = meta_true_d %>%
  filter(meta_task=="toggle_losers",
         base_task_toe == "example_is_train" | source == "game_straight_flush_l_0_bv_0_sr_0") %>%
  group_by(run_type, run, epoch, base_task_toe) %>%
  summarize(mean_loss=mean(loss)) %>%
  ungroup() %>%
  spread(base_task_toe, mean_loss) %>%
  group_by(run_type, run) %>%
  filter(example_is_train == max(example_is_train))
selected_meta_true_d %>%
  arrange(desc(example_is_train))

```