### 2019-12-05 writing/poster/HoMM_transferable_skills_poster.tex
Starting poster

\documentclass[final]{beamer}
\usepackage[size=a0]{beamerposter}
\mode<presentation>{\usetheme{lampinen}}
\usepackage{float}
\usepackage{url}           
%\usepackage{booktabs}      
\usepackage{amsfonts}      
%\usepackage{blkarray}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{nicefrac}      
%\usepackage{microtype}   
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage[numbers]{natbib}
\usetikzlibrary{matrix,backgrounds}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}

\tikzstyle{netnode} = [circle, draw, thick, inner sep=0pt, minimum size=0.5cm]
\tikzstyle{relunode} = [rectangle, draw, thick, inner sep=0pt, minimum size=0.5cm]

\tikzstyle{line} = [draw, line width=0.5pt, -latex']

\newcommand{\R}{\mathbb{R}}
\newcommand{\bb}[1]{{\bf\overline{#1}}}
\newcommand{\bh}[1]{{\bf\hat{#1}}}
\newcommand{\sg}[1]{{\color{red} #1}}
\newcommand{\eopt}[1]{\varepsilon^\text{opt}_{#1}}

\newcommand{\trainerr}{\mathcal{\varepsilon}_\text{train}}
\newcommand{\generr}{\mathcal{\varepsilon}_\text{test}}

\newcommand{\toptn}{{t}^\text{opt}_\text{neural}}
\newcommand{\eoptn}{\mathcal{\varepsilon}^\text{opt}_\text{neural}}
\newcommand{\eoptnn}{\mathcal{\varepsilon}^\text{opt}_\text{non-neural}}

\newcommand{\wa}{{\bf{W}^{21}}}
\newcommand{\wb}{{\bf{W}^{32}}}
\newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ovn}{\overline{N}}

\setlength{\parskip}{0.5em}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\title{Zero-Shot Task Adaptation by Homoiconic Meta-mapping (HoMM)}
\author{Andrew K. Lampinen \& James L. McClelland}
\institute{Department of Psychology, Stanford University}

\begin{document}

\begin{frame}[t]{}
\vspace{-1.5em}
\begin{columns}

\begin{column}[t]{0.248\textwidth}
\begin{block}{\huge Zero-shot adaptation}
\end{block}
\begin{block}{\huge Meta-mapping}
\end{block}
\end{column}
\begin{column}[t]{0.496\textwidth}
\begin{block}{\huge Homoiconic Meta-Mapping (HoMM) architectures}
\end{block}
\end{column}
\begin{column}[t]{0.248\textwidth}
\begin{block}{\huge Conclusions}
\end{block}
\end{column}
\end{columns}
\end{frame}
\end{document}