### 2020-03-30 cards_model_analysis.Rmd
Metaclass lesion

---
title: "Cards model analysis"
author: "Andrew Lampinen"
output: html_document
---

```{r}
library(tidyverse)
```

# data loading

```{r}
parent_dir = "./presentable_results"
subdirs = c("exact_eval", "exact_eval/metaclass_lesion")
language_subdirs = c("exact_eval/language", "exact_eval/tcnh_more_complex/language")

num_runs = 5
```

```{r}
read_config = function(config_file) { 
  config = read_delim(config_file, delim="\n") %>%
    separate(`key, value`, c("key", "value"), sep=",", extra="merge") %>%
    spread(key, value) %>%
    mutate_at(c("base_train_tasks", "base_eval_tasks", "meta_class_train_tasks", "meta_class_eval_tasks", "meta_map_train_tasks", "meta_map_eval_tasks"), function(x) {
      x = gsub("\\\"|[][]| |\'", "", x)
      return(str_split(x, ","))
    } )
}
```

```{r}
load_d = function(results_dir, result_subdirs, num_runs, file_type) {
  d = replicate(num_runs * length(result_subdirs), data.frame())
  index = 1
  for (run_i in 0:(num_runs-1)) {
    for (result_subdir in result_subdirs) {
      filename = sprintf("%s/%s/run%i_%s.csv", results_dir, result_subdir, run_i, file_type)
      print(filename)
      if (!file.exists(filename)) {
        print(paste("skipping ", filename, sep=""))
        next
      }
      if (grepl("config", file_type)) {
        this_d = read_config(filename)
      } else {
        this_d = read.csv(filename, check.names=F, header=T) 
        names(this_d) <- make.unique(names(this_d))

      }
      this_d = this_d %>%
        mutate(run = run_i,
               run_type = result_subdir)
      d[[index]] = this_d
      index = index + 1
    }
  }
  d = bind_rows(d)
  return(d)
}
```

```{r}
config_d = load_d(parent_dir, subdirs, num_runs, "run_config")
loss_d = load_d(parent_dir, subdirs, num_runs, "losses")
lang_d = load_d(parent_dir, language_subdirs, num_runs, "language_losses")
meta_true_d = load_d(parent_dir, subdirs, num_runs, "meta_true_losses")
```

# some manipulation

```{r}
loss_d = loss_d %>%
  filter(epoch %% 100 == 0) %>%
  gather(task, loss, -epoch, -run, -run_type) %>%
  mutate(meta = grepl("is|toggle", task)) %>%
  filter(!meta, grepl("rewards", task)) %>%
  mutate(task = sub("_expected_rewards", "", task),
         train_or_eval = ifelse(grepl("straight_flush_l_1", task), "eval", "train"))

```

```{r}
lang_d = lang_d %>%
  filter(epoch %% 100 == 0) %>%
  gather(task, loss, -epoch, -run, -run_type) %>%
  separate(task, c("task", "train_or_eval"), ":") %>%
  mutate(task = sub("_expected_rewards", "", task))

```

```{r}
meta_true_d = meta_true_d %>%
  filter(epoch %% 100 == 0) %>%
  gather(task, loss, -epoch, -run, -run_type) %>%
  separate(task, c("meta_task", "mapping_toe", "base_task_toe", "source", "target"), ":|->")

```

# basic plots
```{r}
theme_set(theme_bw())
```

```{r}
ggplot(loss_d,
       aes(x=epoch, y=loss, color=train_or_eval)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run_type ~ meta)
```
```{r}
ggplot(loss_d %>%
         filter(task == "game_straight_flush_l_0_bv_0_sr_0:train"),
       aes(x=epoch, y=loss, color=train_or_eval)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run_type ~ run)
```

```{r}
ggplot(meta_true_d %>% 
         filter(meta_task == "toggle_losers"),
       aes(x=epoch, y=loss, color=base_task_toe)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run ~ run_type)
```

```{r}
ggplot(meta_true_d %>% 
         filter(meta_task == "toggle_losers",
                source == "game_straight_flush_l_0_bv_0_sr_0"),
       aes(x=epoch, y=loss, color=base_task_toe)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run ~ run_type)
```

```{r}
ggplot(meta_true_d %>%
         filter(meta_task=="toggle_losers") %>%
         group_by(run_type, run, epoch, base_task_toe) %>%
         summarize(mean_loss=mean(loss)) %>%
         ungroup() %>%
         spread(base_task_toe, mean_loss),
       aes(x=example_is_train, y=example_is_eval, alpha=epoch)) +
  geom_point() +
  facet_grid(run ~ run_type)
```

This seems to be a good criteria for persistent reps, but interestingly not for non-persistent.
```{r}
selected_meta_true_d = meta_true_d %>%
  filter(meta_task=="toggle_losers",
         base_task_toe == "example_is_train" | source == "game_straight_flush_l_0_bv_0_sr_0") %>%
  group_by(run_type, run, epoch, base_task_toe) %>%
  summarize(mean_loss=mean(loss)) %>%
  ungroup() %>%
  spread(base_task_toe, mean_loss) %>%
  group_by(run_type, run) %>%
  filter(example_is_train == max(example_is_train))
selected_meta_true_d %>%
  arrange(desc(example_is_train))

```

This doesn't work particularly well either for non-persistent, either selecting on eval_other or on train_and_other, but might work better with persistent
```{r}
selected_meta_true_d_2 = meta_true_d %>%
  filter(meta_task=="toggle_losers") %>%
  mutate(source_type=case_when(
    base_task_toe == "example_is_train" ~ "train",
    source == "game_straight_flush_l_0_bv_0_sr_0" ~ "eval_target",
    T ~ "eval_other")) %>%
  group_by(run_type, run, epoch, source_type) %>%
  summarize(mean_loss=mean(loss)) %>%
  ungroup() %>%
  spread(source_type, mean_loss) %>%
  mutate(train_and_other = train + eval_other) %>%
  group_by(run_type, run) %>%
  filter(train_and_other == max(train_and_other))
selected_meta_true_d_2 %>%
  arrange(desc(train_and_other))
```

Also somewhat promising:
```{r}
source_task_loss_d = loss_d %>%
  filter(task == "game_straight_flush_l_0_bv_0_sr_0:train") %>%
  select(epoch, run, run_type, loss) %>%
  rename(source_task_base_loss = loss)

selected_meta_true_d_3 = meta_true_d %>%
  filter(epoch > 5000, epoch < 50000) %>%
  filter(meta_task=="toggle_losers") %>%
  mutate(source_type=case_when(
    base_task_toe == "example_is_train" ~ "train",
    source == "game_straight_flush_l_0_bv_0_sr_0" ~ "eval_target",
    T ~ "eval_other")) %>%
  group_by(run_type, run, epoch, source_type) %>%
  summarize(mean_loss=mean(loss)) %>%
  ungroup() %>%
  spread(source_type, mean_loss) %>%
  left_join(source_task_loss_d) %>%
  mutate(source_loss_and_other = source_task_base_loss + eval_other) %>%
  group_by(run_type, run) %>%
  filter(source_loss_and_other == max(source_loss_and_other)) %>%
  slice(1) # take first where multiple minima
selected_meta_true_d_3 %>%
  arrange(desc(source_loss_and_other))
```

```{r}
# window_size = 1000
# 
# sliding_window_selected_meta_true_d = meta_true_d %>%
#   filter(epoch < 50000) %>%
#   filter(meta_task=="toggle_losers") %>%
#   mutate(source_type=case_when(
#     base_task_toe == "example_is_train" ~ "train",
#     source == "game_straight_flush_l_0_bv_0_sr_0" ~ "eval_target",
#     T ~ "eval_other")) %>%
#   group_by(run_type, run, epoch, source_type) %>%
#   summarize(mean_loss=mean(loss)) %>%
#   ungroup() %>%
#   spread(source_type, mean_loss) %>%
#   left_join(source_task_loss_d) %>%
#   mutate(source_loss_and_other = source_task_base_loss + eval_other) %>%
#   mutate(window = epoch %/% window_size) %>%
#   group_by(run_type, run, window) %>%
#   summarize(source_loss = mean(source_task_base_loss), eval_other = mean(eval_other), source_loss_and_other = mean(source_loss_and_other), eval_target = mean(eval_target)) %>%
#   ungroup() %>%
#   group_by(run_type, run) %>%
#   filter(source_loss_and_other == max(source_loss_and_other))
# sliding_window_selected_meta_true_d %>%
#   arrange(desc(source_loss_and_other))
```

# language baselines

```{r}
ggplot(lang_d,
       aes(x=epoch, y=loss, color=train_or_eval)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(run_type ~ run)
```

Equivalent selection for the language baselines to that used above:
```{r}
selected_lang_d = lang_d %>%
  filter(epoch > 5000,
         epoch < 50000) %>%
  mutate(task_type=case_when(
    task == "game_straight_flush_l_0_bv_0_sr_0" ~ "source",
    task == "game_straight_flush_l_1_bv_0_sr_0" ~ "eval_target",
    train_or_eval == "eval" ~ "eval_other",
    T ~ "train_other")) %>%
  group_by(run_type, run, epoch, task_type) %>%
  summarize(mean_loss=mean(loss)) %>%
  ungroup() %>%
  spread(task_type, mean_loss) %>%
  mutate(source_loss_and_other = source + eval_other) %>%
  group_by(run_type, run) %>%
  filter(source_loss_and_other == max(source_loss_and_other)) %>%
  slice(1) # take first where multiple minima
selected_lang_d %>%
  arrange(desc(source_loss_and_other))
```

# data for figure

```{r}
diss_figure_data = bind_rows(
  selected_lang_d %>%
    filter(run_type == "exact_eval/language") %>%
    ungroup() %>%
    select(run, source, eval_target) %>%
    mutate(result_type = "Language baseline"),
  selected_lang_d %>%
    filter(run_type == "exact_eval/tcnh_more_complex/language") %>%
    ungroup() %>%
    select(run, source, eval_target) %>%
    mutate(result_type = "Language baseline\n(task-conditioned)"),
  selected_meta_true_d_3 %>%
    ungroup() %>%
    filter(run_type == "exact_eval/metaclass_lesion") %>%
    rename(source=source_task_base_loss) %>%
    select(run, source, eval_target) %>%
    mutate(result_type = "Meta-class lesion"),
  selected_meta_true_d_3 %>%
    ungroup() %>%
    filter(run_type == "exact_eval") %>%
    rename(source=source_task_base_loss) %>%
    select(run, source, eval_target) %>%
    mutate(result_type = "HoMM")) %>%
  gather(task, expected_reward, source, eval_target)
```

```{r}
write_csv(diss_figure_data, "~/Documents/grad/res/psych/cards/analysis/HoMM_data.csv")
```